---
title: "Application to simulated data"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  pdf_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Application to simulated data}
  %\usepackage[utf8]{inputenc}
---

# Constant population size

Skyline, skyride and skygrid models have a tendency to erroneously identify exponential growth in situations where the size is constant. 
We can ameliorate that error by selecting a smaller time-axis resolution.

Simulate the tree with constant Ne: 
```{r}
library(ape)
library(mlesky)
set.seed(0)
tree=rcoal(100)
tree$edge.length=tree$edge.length*10#So that Ne(t)=10
plot(tree, show.tip.label = F)
axisPhylo(backward = T)
mtext("Years since sampling", side=1, line=3)
```

Find the best `res`:
```{r}
resx = optim_res_aic( tree , ncpu = 8,adapt_time_axis = T)
print(resx)
```
Note that this tests multiple `res` values using multi-core processing. The `ncpu` parameter should be chosen for your system. 

Estimate Ne(t):
```{r}
fit <- mlskygrid( tree, res = resx, tau=NULL, tau_lower = 1,tau_upper = 10000,adapt_time_axis = T)
plot( fit , logy=FALSE)
```

# Varying population size

Let's consider that the effective population size was of size 10 except between 2000 and 2010 
when it was of size 1 (bottleneck model). Ten samples per year are taken at uniform timepoints 
between 1990 and 2020. We can simulate such a phylogeny using:

```{r}
library(ape)
library(mlesky)
set.seed(0)
alphaFun=function(x){if (x<2000 || x>2010) 10 else 1}
sampleDates=seq(1990,2020,0.1)
t=simCoal(sampleDates,alphaFun)
plotBoth(t,alphaFun)
```

We now infer the past population size given the phylogeny, with res selected using the AIC criterion:

```{r}
res=optim_res_aic(t,ncpu = 2)
print(res)
fit=mlskygrid(t,res=res,tau=NULL,tau_lower = 1,tau_upper = 20,ncpu=2)
plot(fit)
```

We now infer the past population size given the phylogeny, with res selected using the BIC criterion:

```{r}
res=optim_res_bic(t,ncpu = 2)
print(res)
fit=mlskygrid(t,res=res,tau=NULL,tau_lower = 1,tau_upper = 20,ncpu=2)
plot(fit)
```
