---
title: "Application to simulated data"
author: "Xavier Didelot"
date: '`r Sys.Date()`'
output:
  pdf_document: default
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Application to simulated data}
  %\usepackage[utf8]{inputenc}
---

# Simulation

Let's consider that the effective population size was of size 10 except between 2000 and 2010 
when it was of size 1 (bottleneck model). Ten samples per year are taken at uniform timepoints 
between 1990 and 2020. We can simulate such a phylogeny using:

```{r}
library(ape)
library(mlesky)
set.seed(0)
alphaFun=function(x){if (x<2000 || x>2010) 10 else 1}
sampleDates=seq(1990,2020,0.1)
t=simCoal(sampleDates,alphaFun)
plotBoth(t,alphaFun)
```

# Inference using AIC

We now infer the past population size given the phylogeny, with res selected using the AIC criterion:

```{r}
res=optim_res_aic(t,ncpu = 6,adapt_time_axis=F)
print(res)
fit=mlskygrid(t,res=res,adapt_time_axis=F,tau=NULL,tau_lower = 1,tau_upper = 20,ncpu=6)
plot(fit)
```

# Inference using BIC

We now infer the past population size given the phylogeny, with res selected using the BIC criterion:

```{r}
res=optim_res_bic(t,ncpu = 6,adapt_time_axis=F)
print(res)
fit=mlskygrid(t,res=res,adapt_time_axis=F,tau=NULL,tau_lower = 1,tau_upper = 20,ncpu=6)
plot(fit)
```
